---
title: "README"
author: "JOSEPH"
date: "`r Sys.Date()`"
output: github_document
---
```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(gtsummary)
library(labelled)
library(ggplot2)

```

# This project examines the relationship between poverty levels and educational attainment across counties in Oklahoma. Specifically, it explores how high school and bachelor's degree completion rates are associated with poverty levels.

The analysis includes:

A summary table comparing education levels across poverty groups
Scatter plots visualizing relationships between poverty and education
A regression model assessing these associations


```{r, load data, echo = FALSE}
data <- readRDS("merged_clean.RDS")
#sellect columns Name, pct_population_below_poverty, pct_adults_25plus_hs_or_higher, pct_adults_25plus_bachelors_or_higher
project_data <- data |> 
                select(NAME, pct_population_below_poverty, pct_adults_25plus_hs_or_higher, pct_adults_25plus_bachelors_or_higher)

```



```{r}
labelled::var_label(project_data) <- list(
  NAME = "County",
  pct_population_below_poverty = "Poverty (%)",
  pct_adults_25plus_hs_or_higher = "High School (%)",
  pct_adults_25plus_bachelors_or_higher = "Bachelors (%)"
)
```


#poverty Groups
```{r}
project_data <- project_data |>
  mutate(
    poverty_group = ifelse(
      pct_population_below_poverty >= median(pct_population_below_poverty, na.rm = TRUE),
      "High Poverty",
      "Low Poverty"
    )
  )
```

#Table 1
```{r}


table1 <- project_data |>
  select(
    poverty_group,
    pct_adults_25plus_hs_or_higher,
    pct_adults_25plus_bachelors_or_higher
  ) |>
  tbl_summary(
    by = poverty_group,
    statistic = all_continuous() ~ "{mean} ({sd})",
    label = list(
      pct_adults_25plus_hs_or_higher ~ "High School Education (%)",
      pct_adults_25plus_bachelors_or_higher ~ "Bachelor's Degree (%)"
    )
  ) |>
  add_p() |>
  modify_header(label ~ "**Variable**") |>
  modify_caption("**The Relationship Between Poverty Levels and Educational Attainment in Oklahoma**")

table1
```

```{r}

# Poverty vs High School
ggplot(project_data, aes(x = pct_adults_25plus_hs_or_higher,
                         y = pct_population_below_poverty)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Poverty vs High School Education",
       x = "High School Education (%)",
       y = "Poverty (%)")

# Poverty vs Bachelors
ggplot(project_data, aes(x = pct_adults_25plus_bachelors_or_higher,
                         y = pct_population_below_poverty)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Poverty vs Bachelor's Education",
       x = "Bachelor's Degree (%)",
       y = "Poverty (%)")
```

#A linear regression model estimates the relationship between poverty and education levels.


```{r}

model <- lm(
  pct_population_below_poverty ~ 
    pct_adults_25plus_hs_or_higher +
    pct_adults_25plus_bachelors_or_higher,
  data = project_data
)

reg_table <- tbl_regression(model) |>
  add_global_p() |>
  modify_caption("**Association Between Poverty Levels and Educational Attainment in Oklahoma**")

reg_table
```

